---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  out.width = "100%"
)
```

# ped suite

<!-- badges: start -->
<!-- badges: end -->

The **ped suite** package is a collection of packages for pedigree analysis in R.

The core packages are:

* [pedtools](https://github.com/magnusdv/pedtools): Creating and working with pedigrees and marker data
* [ribd](https://github.com/magnusdv/ribd): Computation of pedigree-based relatedness coefficients
* [pedprobr](https://github.com/magnusdv/pedprobr): Marker probabilities and pedigree likelihoods
* [pedmut](https://github.com/magnusdv/pedmut): Mutation models for pedigree likelihood computations
* [forrel](https://github.com/magnusdv/forrel): Forensic pedigree analysis and relatedness analysis

In addition, the following packages are devoted to special applications:

* [ibdsim2](https://github.com/magnusdv/ibdsim2): Simulation of identity-by-descent sharing by family members
* [pedbuildr](https://github.com/magnusdv/pedbuildr): Pedigree reconstruction
* [paramlink2](https://github.com/magnusdv/paramlink2): Parametric linkage analysis
* [segregatr](https://github.com/magnusdv/segregatr): Segregation analysis for clinical variant interpretation
* [dvir](https://github.com/thoree/dvir): Disaster victim identification




## Installation

You can install **pedsuite** from [GitHub](https://github.com/magnusdv/pedsuite) with:

``` r
# install.packages("devtools")
devtools::install_github("magnusdv/pedsuite")
```
## Example

```{r}
library(pedsuite)
```

We now have access to all the functions of **pedtools** and the other `ped suite` packages.

```{r ex-ped, fig.height = 3, fig.width = 3.5, out.width = "40%"}
# Create pedigree
x = halfSibPed(sex1 = 1, sex2 = 2)
x = addChildren(x, father = 4, mother = 5)

# Plot it
plot(x, hatched = 6, margins = c(1,1,1,1))
```

Calculate the inbreeding coefficient of the child (with `ribd::inbreeding()`):

```{r}
inbreeding(x, ids = 6)
```

Simulate genotypes for a single marker with alleles `a`, `b`, `c`, `d` (with `forrel::markerSim()`):
```{r}
y = markerSim(x, N = 1, alleles = c("a", "b", "c", "d"), seed = 1729)
```

And plot the result:
```{r ex-ped-2, fig.height = 3, fig.width = 3.5, out.width = "40%"}
plot(y, marker = 1, margins = c(1,1,1,1))
```

