---
title: ISFG Summer School 2023
subtitle: "Workshops 4.1 and 4.2: Pedigree analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{ISFG Summer School 2023}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

* Dates: August 30 & 31, 2023
* Where: Online, as part of [ISFG Summer School 2023](https://isfgsummerschool.idloom.events/isfg-summer-school-virtual-edition-2023)
* Teachers: Magnus Dehli Vigeland and Thore Egeland

## Description
The study of genetic relatedness and the flow of genetic information within families, is central in forensic genetics. The aim of this course is to introduce the elegant statistical foundations of relatedness, and to explore important forensic applications. We use the software R for calculations and visualisations, in particular the [pedsuite](https://magnusdv.github.io/pedsuite/) packages for pedigree analysis. Most lectures and exercises are based on the book [Pedigree Analysis in R](https://www.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2) (but you do not need the book in order to attend the course).

## Learning outcome
After completing the course the participants will have knowledge about:

* Classical concepts of statistical genetics and relatedness analysis
* How to produce informative and beautiful pedigree plots
* How to compute coefficients of relatedness
* How to perform kinship testing and pedigree reconstruction in R
* How to perform disaster victim identification in R

## Prerequisites
* Basic knowledge of genetics and probability
* Some previous experience with R is recommended, but not strictly required
* A recent version of [R](https://www.r-project.org) (at least 4.1) must be installed on your computer
* We recommend running R from [RStudio](https://www.rstudio.com/products/rstudio/download/)

The following R command installs the **pedsuite** packages, which will be used throughout the course:

```{r, eval = F}
install.packages("pedsuite")
```

## Schedule
The course runs from 18 to 22 (CEST) each day, with a 15 minute break in the middle. The following schedule is tentative:

#### Aug 30 (Wednesday) -- Pedigree analysis: Basic
* 18:00--19:00 Lecture 1. [**Pedigrees and measures of relatedness**](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2023/Lecture1_pedigrees.pdf) (MDV)
* 19:00--19:15 Crash course in R (MDV)
* 19:15--20:00 [Exercise set 1](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2023/exercises1.pdf)
* 20:00--20:15 _Break_
* 20:15--21:00 Lecture 2. [**Kinship testing**](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2023/Lecture2_kinship.pdf) (TE)
* 21:00--21:45 [Exercise set 2](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2023/exercises2.pdf)
* 21:45--22:00 Summary and discussion

#### Aug 31 (Thursday) -- Pedigree analysis: Advanced
* 18:00--19:00 Lecture 3. [**Relatedness inference and pedigree reconstruction**](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2023/Lecture3_reconstruction.pdf) (MDV)
* 19:00--19:45 Exercise set 3
* 19:45--20:00 _Break_
* 20:00--21:00 Lecture 4. **Disaster victim identification** (TE)
* 21:00--21:45 Exercise set 4
* 21:45--22:00 Summary and discussion



Comments and questions may be sent to *magnusdv at gmail dot com*.
