---
title: ISFG Summer School - Virtual Edition 2021
subtitle: "Workshop 7: Pedigree analysis in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{ISFG Summer School - Virtual Edition 2021}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

* Title: Pedigree analysis in R
* Dates: July 28-30, 2021
* Where: Online, as part of [ISFG Summer School 2021](https://isfgsummerschool.events.idloom.com/virtual-edition-2021)
* Teachers: Magnus Dehli Vigeland and Thore Egeland

## Description
The study of genetic relatedness and how genetic information flows in families, is central in a variety of research fields, including medical and forensic genetics. The aim of this course is to introduce the elegant statistical foundations of relatedness, as well as several forensic applications, with a particular focus on using R for calculations and visualisations. Most lectures and exercises are based on the book [Pedigree Analysis in R](https://www.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2). A 30% discount will be available for participants (but you do not have to buy the book in order to attend the course).

## Learning outcome
After completing the course the participants should have knowledge about:

* Classical concepts and methods of statistical genetics and relatedness analysis
* How to create, manipulate and plot pedigrees in R
* How to compute pedigree likelihoods and likelihood ratios in R 
* How to perform kinship testing and pedigree reconstruction in R
* How to perform disaster victim identification in R

## Prerequisites
 
 * Basic knowledge of genetics and probability. 
 * Some previous experience with R is recommended.
 * Please make sure that [R](www.r-project.org) (version > 4.0) is installed on your computer. We recommend running R from [RStudio](https://www.rstudio.com/products/rstudio/download/).

The following R command installs the **ped suite** packages, which will be used throughout the course:

```{r, eval = F}
install.packages("pedsuite")
```

## Schedule
The course runs from 14 to 18 (CEST) each day, with a 15 minute break in the middle. The lectures are aimed at 55 minutes, allowing for a short interval before exercises.

#### Day 1 — July 28 (Wednesday)
* 14:00–15:00 Lecture 1. [Introductions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/Lecture1_intro.pdf) (MDV)
* 15:00–15:45 [Exercise set 1](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises1.pdf). ([Solutions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises1-solutions.pdf).)
* 15:45–16:00 Break
* 16:00–17:00 Lecture 2. [Working with pedigrees in R](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/Lecture2_pedsuite.pdf) (MDV)
* 17:00–18:00 [Exercise set 2](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises2.pdf). ([Solutions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises2-solutions.pdf).)

#### Day 2 — July 29 (Thursday)
* 14:00–15:00 Lecture 3. [Kinship testing: paternity cases and complex cases](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/Lecture3-kinship-LR.pdf) (TE)
* 15:00–15:45 [Exercise set 3](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises3.pdf). ([Solutions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises3-solutions.pdf).)
* 15:45–16:00 Break
* 16:00–17:00 Lecture 4. [Relatedness coefficients](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/Lecture4_relatednessI.pdf) (MDV)
* 17:00–18:00 [Exercise set 4](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises4.pdf). ([Solutions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises4-solutions.pdf).)

#### Day 3 — July 30 (Friday)
* 14:00–15:00 Lecture 5. [Relatedness inference and pedigree reconstruction](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/Lecture5-relatedness-inference.pdf) (MDV)
* 15:00–15:45 [Exercise set 5](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises5.pdf). ([Solutions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises5-solutions.pdf).)
* 15:45–16:00 Break
* 16:00–17:00 Lecture 6. [Disaster victim identification and other forensic applications](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/Lecture6-dvi.pdf) (TE)
* 17:00–18:00 [Exercise set 6](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises6.pdf). ([Solutions](https://github.com/magnusdv/pedsuite/blob/master/vignettes/web_only/courses/isfg2021/exercises6-solutions.pdf).)
  
  
Comments and questions may be sent to *magnusdv at gmail dot com*.
