---
title: "Workshop: Pedigrees and kinship analysis in forensic genetics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Workshop: Pedigrees and kinship analysis in forensic genetics}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```


* Date: April 28--29, 2026
* Where: Alcalá de Henares, Spain
* Teachers: Magnus Dehli Vigeland & Thore Egeland
* Organiser: Instituto Universitario de Investigación en Ciencias Policiales (IUICP)
* Official [home page](https://iuicp.uah.es/es/actividades/congresos/Pedigrees-and-kinship-analysis-in-forensic-genetics/)

## Description
The study of pedigrees and genetic relatedness is central in forensic genetics. The
aim of this workshop is to introduce the statistical foundations of relatedness, 
and explore several forensic applications. In particular we focus on recent user-friendly apps,
that make pedigree analysis more accessible to practitioners. These include QuickPed, KLINK and Diviana, which are all available as online tools.

Parts of the course is based on the book [Pedigree Analysis in R](https://www.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2) (Vigeland '21), which presents the 
pedsuite R packages underlying the mentioned apps. 

We also demonstrate the [Familias](https://familias.no) software, which is widely used for forensic kinship testing and disaster victim identification (DVI).

Each day alternates between lectures and hands-on exercises. Detailed solutions for all exercises are provided at the end of each day, and discussed as needed.


## Learning outcome
After completing the course the participants will know:

* Classical concepts of statistical genetics and relatedness analysis
* How to create and analyse pedigrees with QuickPed
* How to perform kinship testing in Familias and KLINK
* How to perform disaster victim identification with Diviana

## Prerequisites

* All participants should bring a laptop on which they have administrative rights
* Recent versions of [R](https://www.r-project.org) and [RStudio](https://www.rstudio.com/products/rstudio/download/) should be installed.
* In addition, you should install a recent version of [Familias](https://familias.no/download).

If you run into installation problems, we will help during the workshop, but to save time for everybody, please try to install these in advance.

## Schedule

The workshop is run over two days, April 28--29, 2026.

```{r, echo = F}
link = pedsuite:::linkFUN(folder = "courses/alcala2026")
```

### Day 1 (Tuesday, April 28) -- Relatedness and kinship analysis

* 08:30--08:45 Registration
* 08:45--09:00 Introductions
* 09:00--10:00 `r link("Pedigrees and measures of relatedness", "Lecture1_pedigrees.pdf")` (MDV)
* 10:00--11:00 `r link("Exercises I", "exercises1.pdf")`. (*Software: QuickPed*)
<!-- -->
<li style="list-style:none; color:tomato;">11:00--11:30 COFFEE BREAK</li>
<!-- -->
* 11:30--12:30 `r link("Forensic kinship testing", "Lecture2_kinship.pdf")` (TE)
* 12:30--13:30 `r link("Exercises II", "exercises2.pdf")`. (*Software: Familias + R*)
<!-- -->
<li style="list-style:none; color:tomato;"><strong>13:30--15:00 LUNCH</strong></li>
<!-- -->
* 15:00--16:00 `r link("Kinship testing with linked markers", "Lecture3_linked.pdf")` (MDV)
* 16:00--17:00 `r link("Exercises III", "exercises3.pdf")`. (*Software: KLINK*)

### Day 2 (Wednesday, April 29) -- Identification cases

* 09:00--10:30 `r link("Disaster victim identification", "Lecture4_dvi.pdf")` (TE)
<!-- -->
<li style="list-style:none; color:tomato;">10:30--11:00 COFFEE BREAK</li>
<!-- -->
* 11:00--12:00 `r link("Exercises IV", "exercises4.pdf")`. (*Software: DIVIANA*)
* 12:00--13:30 `r link("Realised relatedness and IBD segments", "Lecture5_realised.pdf")` (MDV)
<!-- -->
<li style="list-style:none; color:tomato;"><strong>13:30--15:00 LUNCH</strong></li>
<!-- -->
* 15:00--16:00 `r link("Exercises V", "exercises5.pdf")`. (*Software: ibdsim2*)
* 16:00--17:00 `r link("Case study: Identifying a victim of the Marcinelle mining disaster", "Lecture6_marcinelle.pdf")`


Solutions to the exercises can be requested by email to *magnusdv at gmail dot com*.
