---
title: "GHEP Online School 2024: Workshop 1"
subtitle: "Kinship and pedigree analysis: Methods and applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{GHEP Online School 2024: Workshop 1}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

* __Dates__: October 7, 14, 21 and 28, 2024
* __Where__: Online, as part of [ISFG-GHEP Online School 2024](https://ghep-isfg.org/en/workshops/ghep-os/autumn2024/WS1/)
* __Teachers__: Magnus D. Vigeland and Thore Egeland
* __Audience__: Students, researchers and case workers in forensic genetics

## Learning outcomes
* Classical concepts and methods of genetic relatedness 
* Using QuickPed and other apps for pedigree analysis
* Using Familias software to perform kinship testing
* Methods for relationship inference and pedigree reconstruction
* Practical methods for disaster victim identification

## Software preparation: R
Several exercises will use R. In order to do them, please install the following on your computer:

* A recent version of [R](https://www.r-project.org)
* [RStudio](https://www.rstudio.com/products/rstudio/download/)

Run the following command in R to install the **pedsuite** packages:

```{r, eval = F}
install.packages("pedsuite")
```

## Software preparation: Familias
Part of the course is based on kinship testing in the Familias software, which can be installed here: https://familias.no/download.

```{r, echo = F}
courseFolder = "courses/ghep2024"

# Function for creating links to files
link = function(tit, file = NULL, bold = F, live = T) {
  if(bold)
    tit = sprintf("**%s**", tit)
  
  if(!live)
    return(tit)

  locpth = file.path(courseFolder, file)
  
  if(!isTRUE(file.exists(locpth))) {
    cat("File does not exist locally:", locpth, "\n")
    return(tit)
  }
  
  # Formatted link
  url = "https://raw.githubusercontent.com/magnusdv/pedsuite/master/vignettes/web_only"
  sprintf('[%s](%s/%s)', tit, url, locpth)
}
```
```

## Schedule
The course runs each Monday of October 2024, from 16 to 20 (CEST). The following schedule is tentative:

#### Oct 7: Theory of relatedness
* 16:00--17:00 `r link("Introduction to pedigrees, QuickPed and R", "ghep2024_lecture1_pedigrees.pdf", bold = T)` (MDV)
* 17:00--17:45 `r link("Exercises", "exercises1.pdf")`. (`r link("Solutions", "solutions1.pdf")`)
* 17:45--18:00 _Break_
* 18:00--19:00 `r link("Measures of relatedness", "ghep2024_lecture2_coefficients.pdf", bold = T)` (MDV)
* 19:00--19:45 `r link("Exercises", "exercises2.pdf")`. (`r link("Solutions", "solutions2.pdf")`)
* 19:45--20:00 Wrap-up

#### Oct 14: Kinship testing
* 16:00--17:00 `r link("Introduction to forensic kinship testing", "ghep2024_lecture3_testing.pdf", bold = T)` (TE)
* 17:00--17:45 `r link("Exercises", "exercises3.pdf")`
* 17:45--18:00 _Break_
* 18:00--19:00 `r link("Kinship testing with Familias", "ghep2024_lecture4_Familias.pdf", bold = T)` (TE)
* 19:00--19:45 `r link("Exercises", "exercises4.pdf")`. File needed: `r link("kinship-riddle.fam", "kinship-riddle.fam")`
* 19:45--20:00 Wrap-up

#### Oct 21: Relatedness inference
* 16:00--17:00 `r link("Realised relatedness: Why are some siblings more alike than others?", "ghep2024_lecture5_realised.pdf", bold = T, live = F)` (MDV)
* 17:00--17:45 `r link("Exercises", "exercises5.pdf", live = F)`
* 17:45--18:00 _Break_
* 18:00--19:00 `r link("Pedigree reconstruction", "ghep2024_lecture6_reconstruction.pdf", bold = T, live = F)` (MDV)
* 19:00--19:45 `r link("Exercises", "exercises6.pdf", live = F)`
* 19:45--20:00 Wrap-up

#### Oct 28: Disaster victim identification
* 16:00--17:00 `r link("DNA-based disaster victim identification", "ghep2024_lecture7_dvi.pdf", bold = T, live = F)` (TE)
* 17:00--17:45 `r link("Exercises", "exercises7.pdf", live = F)`
* 17:45--18:00 _Break_
* 18:00--19:00 `r link("Practical DVI with Diviana", "ghep2024_lecture8_diviana.pdf", bold = T, live = F)` (MDV)
* 19:00--19:45 `r link("Exercises", "exercises8.pdf", live = F)`
* 19:45--20:00 Wrap-up

Comments and questions may be sent to *magnusdv at gmail dot com*.
