---
title: "ISFG 2026: Pre-congress workshop"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{ISFG 2026: Pre-congress workshop}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

# Pedigree analysis and relatedness inference

* Date: 18th of August, 2026
* Where: Montreal, Canada
* Teachers: Magnus Dehli Vigeland & Thore Egeland
* Duration: Full day workshop

## Description
The study of pedigrees and genetic relatedness is central in forensic genetics. The
aim of this course is to introduce the elegant statistical foundations of relatedness, as
well as several forensic applications. In certain parts we use R for calculations and
visualisations. In the basic session we will focus on pedigree coefficients and
likelihood ratios for kinship analysis. In addition, we will demonstrate QuickPed, an
online tool for creating pedigrees and computing relatedness: https://magnusdv.shinyapps.io/quickped

In the advanced part we will explore recently developed tools for Disaster Victim Identification (DVI) and pedigree reconstruction. 

The course material is based on the book [Pedigree Analysis in R](https://www.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2) (Vigeland '21) and selected papers. Each session will alternate between lectures and hands-on
exercises. Detailed solutions for all exercises will be provided and discussed as needed.

Attendance in the ‘Basic session’ is not required for the ‘Advanced session’ for participants who are familiar with R.


## Learning outcome
After completing the course the participants will have knowledge about:

* Classical concepts of statistical genetics and relatedness analysis
* How to produce informative and beautiful pedigree plots
* How to compute coefficients of relatedness
* How to perform kinship testing and pedigree reconstruction in R
* How to perform disaster victim identification in R

## Prerequisites
* Basic knowledge of genetics and probability
* All participants must bring a laptop with administrative rights
* A recent version of [R](https://www.r-project.org) (at least v4.2) must be installed on your computer
* We strongly recommend running R from [RStudio](https://www.rstudio.com/products/rstudio/download/)


## Schedule
The workshop is run as a full-day course on Tuesday, August 18th 2026, from 8:30 to 18:00. The following schedule is tentative:

```{r, echo = F}
link = pedsuite:::linkFUN(folder = "courses/isfg2026")
```

#### Morning session -- Pedigree analysis: Basic
* 08:30--09:30 `r link("Pedigrees and measures of relatedness", "Lecture1_pedigrees.pdf")` (MDV)
* 09:30--10:30 `r link("Exercises I", "exercises1.pdf")`
* 10:30--10:45 _Coffee break_
* 10:45--11:30 `r link("Kinship testing", "Lecture2-Kinship.pdf")` (TE)
* 11:30--12:15 `r link("Exercises II", "exercises2.pdf")`
* 12:15--12:30 Summary and discussion  

<p style="text-indent: 12px;">**Lunch break 12:30 - 14:00**</p>

#### Afternoon session -- Pedigree analysis: Advanced
* 14:00--15:00 `r link("Relatedness inference and pedigree reconstruction", "Lecture3-reconstruction.pdf")` (MDV)
* 15:00--15:45 `r link("Exercises III", "exercises3.pdf")`
* 15:45--16:00 _Coffee break_
* 16:00--17:00 `r link("Disaster victim identification", "Lecture4-DVI.pdf")` (TE)
* 17:00--17:45 `r link("Exercises IV", "exercises4.pdf")`
* 17:45--18:00 Summary and discussion

Solutions to the exercises can be requested by email to *magnusdv at gmail dot com*.
