---
title: Statistical methods in relatedness and pedigree analysis - 2022
subtitle: Arranged by [NORBIS](https://norbis.w.uib.no/), a Norwegian research school in bioinformatics, biostatistics and systems biology.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Statistical methods in relatedness and pedigree analysis - 2022}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

----

## Summary

* Title: Statistical methods in relatedness and pedigree analysis
* Dates: 13--17 June 2022
* Teachers: Magnus Dehli Vigeland and Thore Egeland
* Where: Room 2203a ("Helix"), Kristine Bonnevies hus, University of Oslo
* Address: [Blindernveien 31, entrance from Moltke Moes vei](https://www.uio.no/english/about/getting-around/areas/blindern/bl18/).
* Credits: 4 ECTS
* Exam: Home project (passed/not passed)

## Description
The course introduces statistical methods used in relatedness and pedigree analysis, with applications in forensic genetics, medical genetics and pedigree reconstruction.

The study of genetic relatedness and how genetic information flows in families, is central in a variety of important research fields, including medical genetics, forensics and animal breeding. The aim of this course is to introduce the students to the elegant concepts and results forming the theoretical foundation of genetic relatedness, and to present in some detail several modern applications. In the exercises we use R and the [ped suite](https://magnusdv.github.io/pedsuite/index.html) packages for pedigree analysis and visualisation.

Much of the course material is based on the book [Pedigree Analysis in R](https://www.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2) which may be of interest to some participants (but is not required to attend the course).

## Learning outcome
After completing the course the participants will have knowledge about:

* Classical concepts and methods of statistical genetics and relatedness analysis
* Important applications of pedigree analysis, including forensic kinship analysis, medical segregation analysis and pedigree reconstruction
* How to work with pedigrees in R

## Prerequisites
 
 * Basic knowledge of genetics and probability. 
 * Some previous experience with R is recommended, but not strictly required.
 * Please make sure that [R](www.r-project.org) (version > 4.1) is installed on your computer. We strongly recommend running R from [RStudio](https://www.rstudio.com/products/rstudio/download/).

The following R command installs the **ped suite** packages, which will be used throughout the course:

```{r, eval = F}
install.packages("pedsuite")
```

## Schedule
The course runs from 9 to 16.30 (CEST) each day. A tentative schedule is given below. Minor changes may occur.

#### Monday, June 13

* 09--11 Introduction I: Pedigrees, genetics and probabilities (MDV)
* 11--12 Exercises I 
* 12--13 Lunch
* 13--15 Introduction II: R and the ped suite (MDV)
* 15--16:30 Exercises II

#### Tuesday, June 14

* 09--11 Relatedness I: IBD and coefficients of relatedness (MDV)
* 11--12 Exercises III 
* 12--13 Lunch
* 13--15 Relatedness II: Realised relatedness (MDV)
* 15--16:30 Exercises IV 

#### Wednesday, June 15

* 09--11 Forensic applications I (TE)
* 11--12 Exercises V 
* 12--13 Lunch
* 13--15 Inference of relatedness and pedigree reconstruction (MDV)
* 15--16:30 Exercises VI 

#### Thursday, June 16

* 09--11 Forensic applications II (TE)
* 11--12 Exercises VII
* 12--13 Lunch
* 13--15 Linkage and segregation analysis in medical genetics (MDV)
* 15--16:30 Exercises VIII

#### Friday, June 17

* 09--11 Genealogical triangulation (TE/MDV)
* 11--12 Exercises
* 12--13 Lunch
* 13--16:30 Case study and wrap up (MDV)

<br>

Comments and questions may be sent to *magnusdv at medisin uio no*.
