---
title: Young Investigator Day 2024
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Young Investigator Day 2024}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

# Workshop: Genetic relatedness

* Date: Monday 14th of October, 2024
* Where: Cancer Registry of Norway
* Teacher: Magnus Dehli Vigeland
* Part of the [NOFE conference](https://nofe.no/nofe-conference-2024/) (Norwegian Epidemiological Association)

## Description
The study of pedigrees and genetic relatedness is central in many fields, including medical genetics, forensics and epidemiology.
The aim of this course is to introduce the elegant statistical foundations of relatedness, as well as relevant software tools. 

The workshop will alternate between lectures and hands-on exercises. 
Some exercises use R, while others use _QuickPed_, a free online tool for creating and analysing pedigrees: 
https://magnusdv.shinyapps.io/quickped

The course material is based on the book [Pedigree Analysis in R](https://www.elsevier.com/books/pedigree-analysis-in-r/vigeland/978-0-12-824430-2) (Vigeland '21), 
which may be a point of reference for participants who want to delve deeper into the subject.

## Learning outcome
After completing the course the participants will have knowledge about:

* Classical concepts of statistical genetics and relatedness analysis
* How to produce informative and beautiful pedigree plots
* How to compute coefficients of relatedness
* Why some siblings are more related than others


## Prerequisites
* Basic knowledge of genetics and probability
* All participants must bring a laptop with a recent version of R installed (at least 4.3.0)
* I strongly recommend running R from [RStudio](https://www.rstudio.com/products/rstudio/download/)


## Schedule
The workshop is run as a half-day course, from 12 to 16, with the following (tentative) schedule:

```{r, echo = F}
link = pedsuite:::linkFUN(folder = "courses/yid2024")
```

* 12:00--13:00 `r link("Introduction to pedigrees and relationships", "yid2024_lecture1_pedigrees.pdf")`
* 13:00--13:45 `r link("Exercises I", "exercises1.pdf")`
* 13:45--14:00 _Break_
* 14:00--15:00 `r link("Measures of relatedness", "yid2024_lecture2_coefficients.pdf")`
* 15:00--15:45 `r link("Exercises II", "exercises2.pdf")`
* 15:45--16:00 Summary and discussion  

Solutions to the exercises can be requested by email to *magnusdv at gmail dot com*.

