---
title: QuickPed - a free online pedigree creator
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{QuickPed - a free online pedigree creator}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.align = "center",
                      collapse = TRUE,
                      comment = "#>")
```

::: { .greeting .message style="color: red;"}
**_Try QuickPed here_**: https://magnusdv.shinyapps.io/quickped.
:::

## What is QuickPed?
QuickPed is an interactive web application for drawing and analysing pedigrees. A created pedigree may be saved as an image or as a text file in *ped format* (see [below](#ped-files)). You may also obtain various information about the pedigree, including relatedness coefficients and verbal descriptions of relationships.

QuickPed is powered by the [ped suite](https://magnusdv.github.io/pedsuite/) and [kinship2](https://CRAN.R-project.org/package=kinship2) for pedigree plotting. The web app was built with Shiny.


## Getting started
Creating pedigrees with QuickPed is very intuitive: Select a suitable start pedigree and modify it as needed. You may also load an existing ped file (see below). Modifications are done by clicking on one or several individuals and then applying appropriate buttons, for instance to add children, siblings or parents. At any time you may change attributes like sex, affection status, twin status and ID labels. 

## Tips and tricks

* __Selecting individuals__. Select/deselect pedigree members by clicking on them. Selected individuals are shown in red colour. Time-saving tip: To deselect everyone, click the "Selection" button under the "Remove" heading.

* __Automatic labelling__. Automatic labelling of the pedigree members are available in two different formats. The button marked `1,2,..` applies numeric labels to all individuals, in the order of their appearance in the pedigree plot. Alternatively, the `I-1, I-2,..` button numbers the members generation-wise, using roman numerals to indicate the generation number. 

* __Line breaks in labels__. Double spaces are shown as line breaks in the plot labels. For instance, if you want "King Lear" to appear on two lines, modify the label to "King`r knitr::asis_output("\U2423\U2423")`Lear", i.e., with two spaces between the words.

* __Unknown sex__. Double clicking on a pedigree member changes its symbol into a diamond, representing unknown sex. Double click again to revert. Note: Only pedigree *leaves* (members without children) may have unknown sex.

* __Plot settings__. If the pedigree gets too large, increase the plot region using the control panel on the far right. Here you may also adjust the margins, the size of pedigree symbols and text labels.

## Built-in pedigrees
In the left-most panel of QuickPed the user may choose among a selection of standard pedigrees, including trios, full/half siblings, avuncular and cousin pedigrees of different kinds. Also included are several interesting (albeit less common) pedigree structures like double cousins and quad half first cousins. 

Finally, the following historic pedigrees are available:

* __Habsburg__: A subset of the infamously inbred family tree of the Habsburg royalties. The inbreeding coefficient of King Charles II of Spain (1661-1700) was approximately 0.25, i.e., equivalent to that of a child produced by full siblings. Pedigree adapted from [Wikipedia](https://en.wikipedia.org/wiki/Habsburg_family_tree). See also [The Role of Inbreeding in the Extinction of a European Royal Dynasty](https://doi.org/10.1371/journal.pone.0005174).

* __Jicaque__: A pedigree of Jicaque Indians originally studied by Chapman & Jacquard (1971) and subsequently used in many papers on relatedness and pedigree coefficients.

* __Queen Victoria (haemophilia)__: The royal family tree descending from Queen Victoria, showing the X-linked inheritance of *haemophilia*. Adapted from Figure S1 of [Genotype Analysis Identifies the Cause of the “Royal Disease"](https://doi.org/10.1126/science.1180660).

* __Tutankhamun__: The family tree of the Egyptian pharao Tutankhamun, as inferred from genetic evidence presented by Hawass et al. (2010), [Ancestry and Pathology in King Tutankhamun's Family](https://doi.org/10.1001/jama.2010.121).

## Relationship information
The four buttons in the `Relationships` frame can be used to analyse the relatedness between selected individuals in the current pedigree. They are, in order:

* __Describe relationship__. Prints a textual description of the relationship, generated by [verbalisr](https://github.com/magnusdv/verbalisr). Requires exactly 2 individuals to be selected.
 
* __Calculate coefficients__. Prints a variety of pedigree coefficients between two selected individuals.
  * The inbreeding coefficient of each individual.
  * The kinship coefficient $\varphi$.
  * The degree of relationship, as used e.g. by [KING](https://www.kingrelatedness.com/manual.shtml).
  * The IBD coefficients $\kappa = (\kappa_0, \kappa_1, \kappa_2)$, defined as the probabilities of sharing 0, 1, and 2 alleles identical by descent (IBD). These are included only if both individuals are noninbred (otherwise they are not defined).
  * The 9 condensed identity coefficients of Jacquard, $\Delta = (\Delta_1, ..., \Delta_9)$. These are included only if at least one of the individuals is inbred. (For noninbred individuals, $\Delta = (0,0,0,0,0,0,\kappa_2, \kappa_1, \kappa_0)$.)
    
  More information about these coefficients can be found in the documentation of the [ribd](https://github.com/magnusdv/ribd) package, which is used in the calculations.
  
* __Relatedness triangle__. Produces a plot representing $\kappa$ as a point in a *relatedness triangle*. Requires exactly two noninbred individuals to be selected.

* __Table of coefficients__. Activates a popup window where the user can choose which individuals and coefficients to compute. Generates a tab-separated file for download, containing all the data.


## Ped files
A useful feature of QuickPed is to produce text files describing pedigrees in so-called *ped format*. Such files are often required by software for pedigree analysis.

For a simple illustration, consider this pedigree:

```{r trio-ped, echo = F, fig.height = 1.8, fig.width = 2.2, out.width = "25%", fig.align = "center"}
df = data.frame(id = 1:3, fid = c(0,0,1), mid = c(0,0,2), sex = c(1,2,2), aff = c(1,1,2))
x = pedtools::as.ped(df[1:4])
plot(x, aff = which(df$aff == 2), margins = c(1,1,1,1))
```

A text file describing this pedigree may contain the following.

```{r, echo = F, comment = NA}
print(df, row.names = F)
```

The columns are:

* `id`: Individual ID
* `fid`: Father's ID (or 0 if not included in the pedigree)
* `mid`: Mother's ID (or 0 if not inlcuded in the pedigree)
* `sex`: Sex (1 = male; 2 = female; 0 = unknown)
* `aff`: Affection status (1 = unaffected; 2 = affected; 0 = unknown)

It should be noted that the ped format is not completely standardised, and different software may use slightly different versions. For example, a first column with *Family ID* is sometimes required. Also, the `aff` column may not be needed in non-medical applications. These and other details may be specified when using QuickPed.

Some pedigree information may be shown on the plot, but is *not* stored in the ped file. In the current version of QuickPed, this includes twin relationships, carrier status and deceased status.
